---
title: "NBA players comparision and analysis"
author: "Aneesh Singh"
date: "04/17/2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
geometry: left=0.5in,right=0.5in,top=0.5in,bottom=0.5in
urlcolor: blue
---
Research question - What is the best possible starting 5 and bench 5 based on 2023- 2024 statistics and salary cap?

Where did you find them?
On the NBA stats website
```{r}
library(tidyverse)
library(dcData)
library(ggplot2)
library(dplyr)
nba_data <- read.csv("C:/Users/Singh/Downloads/NBAstats - Sheet1.csv")
nba_salary <- read.csv("C:/Users/Singh/Downloads/Nba Salary Stats - Sheet1.csv")
```
```{r}
glimpse(nba_data)
```

This shows us all the detailed key statistics for each player to help us answer our question.

```{r}
head(nba_salary)
```
This shows us the salary by year of all of the players to help us with our question and find out who is the most effective based on the salary cap. We can see the top players here 

Who collected/maintains them?
The official NBA organization records statistics and publishes them for the public to see. They are updated heavily and are reliable/new 

When & Why were they originally collected?
They were collected during the season in order to see how players perform during every season. This is done to see player/team progression and track data to make better game time decisions such as lineups, defense, shot selections, etc. 

What does a case represent in each data source, and how many total cases are available?
A case in our main data source represents a player and their stats for the season. There are 572 players that played this season. 

What are some of the variables that you plan to use?
We plan to use as many variables as possible, but we primarily will focus on salary, points per game (PPG), rebounds (REB), and assists (AST). For example, we will look at a player's individual statistics in each category and determine who is the best.

Our primary data is nba_data and our secondary is nba_salary
```{r}
nba_data <- read.csv("C:/Users/Singh/Downloads/NBAstats - Sheet1.csv")
nba_salary <- read.csv("C:/Users/Singh/Downloads/Nba Salary Stats - Sheet1.csv")
```

Informative graphs
```{r}
top_25pluspoints <-
  nba_data %>% 
  filter(PTS > 25) 
  
top_25pluspoints %>%
  ggplot(aes(x = PTS, y = reorder(PLAYER, PTS))) +
  geom_point()
 
```


With this plot, we can see two outlier of the player Joel Embiid, and Luka Doncic as they are the leading scorers in the NBA. 

Preliminary question- We might think about putting Joel Embiid and Luka Doncic on out team because of how good of scorers they are. 

Informative graph
```{r}
top_8plusrebounds <-
  nba_data %>% 
  filter(REB > 8) 

top_8plusrebounds %>%
  ggplot(aes(x = REB, y = reorder(PLAYER, REB))) +
  geom_point()
```

With this plot, we can see who the top rebounders are in the entire leauge for this past season. We can seethat Damontas Sabonis sticks out as an outlier.

Preliminary question- Our team is going to need a good rebounders so we are definitely going to try picking someone in the top 10. 

```{r}
top_6plusassists <-
  nba_data %>% 
  filter(AST > 6) 

top_6plusassists %>%
  ggplot(aes(x = AST, y = reorder(PLAYER, AST))) +
  geom_point()
```

With this graph we are able to see who the top players with average assists are. Off the bat, we can see Haliburton and Young as outliers at the top along with Luka Trailing behind.

Preliminary Question- After all of our graphs, we can see Luka Doncic near the top in points, rebounds, and assists so we will definitely favor putting him in our starting 5. Besdies that, there are a lot of players averging high assists so we will consider them for our starting and bench. 


Join two tables together 
```{r}
nba_data <- read.csv("C:/Users/Singh/Downloads/NBAstats - Sheet1.csv")
nba_salary <- read.csv("C:/Users/Singh/Downloads/Nba Salary Stats - Sheet1.csv")

nba_data <- rename(nba_data, Player = PLAYER)

nba_stats <- full_join(nba_data, nba_salary, by = "Player")
head(nba_stats)
```

```{r}
top_blocks <-
  nba_stats %>%
  arrange(X2023.24) %>%
  filter(BLK > 1.2)

top_blocks %>%
  ggplot(aes(x = X2023.24, y = BLK, label = Player)) +
  geom_text(nudge_y = 0.1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_point()
```


```{r}
top_points <-
  nba_stats %>%
  arrange(X2023.24) %>%
  filter(PTS > 25)

top_points %>%
  ggplot(aes(x = X2023.24, y = PTS, label = Player)) +
  geom_text(nudge_y = 0.1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_point()
```


```{r}
top_rebounds <-
  nba_stats %>%
  arrange(X2023.24) %>%
  filter(REB > 9)

top_rebounds %>%
  ggplot(aes(x = X2023.24, y = REB, label = Player)) +
  geom_text(nudge_y = 0.1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_point()
```

```{r}
top_assists <-
  nba_stats %>%
  arrange(X2023.24) %>%
  filter(AST > 7)

top_assists %>%
  ggplot(aes(x = X2023.24, y = AST, label = Player)) +
  geom_text(nudge_y = 0.1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_point()
```
136 million per year salar

Pg: Tyres Haliburton - about 5 million
Sg: Luka Donic - about 40 million
Sf: Anthony Edwards - about 13.5 million
Pf: Giannis - about 45 million
c: Wemby - about 12 million

```{r}
starting_players <- 
  nba_stats %>%
  filter(Player %in% c("Tyrese Haliburton", "Luka Doncic", "Anthony Edwards", "Giannis Antetokounmpo", "Victor Wembanyama")) %>%
  mutate(X2023.24 = as.numeric(str_replace_all(X2023.24, "[$,]", ""))) %>%
  mutate(Player = reorder(Player, -X2023.24))

head(starting_players)
```

```{r}
starting_players %>%
  ggplot(aes(x = Player, y = X2023.24, fill = Player)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "2023-24 best NBA starting 5 in salary cap",
       x = "Player",
       y = "Salary (in millions)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(starting_players, aes(x = "", y = X2023.24)) +
  geom_violin(fill = "lightblue", color = "darkblue") +
  labs(title = "Violin Plot of NBA Starting 5 Salaries for 2023-24",
       y = "Salary ($)",
       x = "") +
  theme_minimal()
```




```{r setupFiles, include = FALSE}
# This template file is based off of a template created by Alex Hayes
# https://github.com/alexpghayes/rmarkdown_homework_template

# Setting Document Options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```


# Code Appendix

```{r codeAppendix, ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}

```